Introduction to LLL\\``Cryptography''

### 2021-06-03 paper.tex
Add introduction and gauss algorithm

\usepackage{alltt, fancyvrb, url}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage{amsmath,stmaryrd,mathtools,algorithm}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{url}

@@

@@

@@

\begin{definition}
    \textbf{Vector space}.
\end{definition}
A \texttt{vector space} $V$ is a subset of $\R^{m}$ which is closed under finite vector addition and scalar multiplication, with the property that

@@

Let $v_1,v_2,\ldots,v_k \in V$. A \texttt{linear combination} of $v_1,v_2,\ldots,v_k \in V$ is any vector of the form

@@

A set of vectors $v_1,v_2,\ldots,v_k \in V$ is \texttt{linearly independent} if the the only way to get

@@

Taken a set of linearly independent vectors $b = (v_1,\ldots,v_n) \in V$ we say that $b$ is a \texttt{basis} of $V$ if $\forall w \in V$ we can write

@@

The vector's length or \texttt{Euclidean norm} of $v = (x_1, x_2, \ldots, x_m)$ is

@@

Let $v, w \in V \subset \R^m$ and $v = (x_1, x_2, \ldots, x_m), w = (y_1, y_2, \ldots, y_m)$, the \texttt{dot product} of $v$ and $m$ is

@@

\begin{figure}[!b]
    \centering
    \includegraphics[scale=0.2]{./img/dot_product.png}
    \caption{Dot Product By 3Blue1Brown}
    \label{fig:dot_product}
\end{figure}

@@

An \texttt{ortoghonal basis} for a vector space $V$ is a basis $v_1, \ldots, v_m$ with the property that

@@

If $\lVert v_i \rVert = 1$ for all $i$ then the basis is \texttt{orthonormal}.

@@

Let $v_1,\ldots,v_n \in \R^m, m \ge n$ be linearly independent vectors. A \texttt{Lattice} $L$ spanned by $\{v_1,\ldots,n_n\}$ is the set of 
all integer linear combinations of $v_1,\ldots,v_n$.

@@

If $v_i$ for every $i = 1,\ldots\,n$ has integer coordinates then the lattice is
called \texttt{Integral Lattice}.

@@

\begin{algorithm}
    \textbf{Gauss Reduction}
\end{algorithm}

@@

\chapter{Cryptanlysis}

@@

\begin{itemize}
    \item \textasciitilde 170 unkown bits of message from an RSA 1024-bit modulus
    \item \textasciitilde 341 unkown bits of message from an RSA 2048-bit modulus
    \item \textasciitilde 683 unkown bits of message from an RSA 4096-bit modulus
\end{itemize}

@@

@@

\subsection{Recover $d$ from many signatures}

@@

\subsection{Recover $d$ knowing MSB bits from each $k$}

>>>>

\usepackage{alltt, fancyvrb, url}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
%\usepackage{algorithmicx}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage{amsmath,stmaryrd,mathtools,algorithm}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{float}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{url}

@@

\chapter{Introduction}

Lattices are powerful mathematical objects that can be used in computer science and mathematics to solves an extensive range of different problems.
In the recents years many post-quantum cryptographic protocols candidates involve the use of hard lattices problems,
for example the \textbf{Learning With Errors} \cite{LWE}.
Lattice constructions can also be used in cryptanalysis to break cryptographic schemes, typically using reduction
algorithms as \textbf{LLL} \cite{LLL}.

To introduce the mathematics of lattices a basic linear algebra background is needed, so the second chapter is a summary of the
basics properties needed to understand it. We used as main resource ``An Introduction to Mathematical Cryptography''\cite{mathcrypto14} to
explain them, but we also tried to gives some geometrically intuition behind some operations.

Cryptographic protocol as \textbf{RSA} and \textbf{ECDSA} are considered secure, however there are various attacks against a
relaxed-model of each of them, i.e., What if we know some bits of the secret key? What if we know a part of the message before the encryption?
We present an introduction to the attacks which involves lattices against these protocols. The code used to confirm the correctness
of the attacks is available on the repository \cite{repo} and the main reference was
``Recovering cryptographic keys from partial information, by example''\cite{cryptoeprint:2020:1506}.

@@

We'll start with some definitions and properties needed to understand the lattices, to move later on the mathematics behind them and some
hard problems related to it. At the end of the chapter the reader should be able to follow the math used in the other chapters.

@@

\begin{definition}
    \textbf{Vector space}.
\end{definition}
A \textit{vector space} $V$ is a subset of $\R^{m}$ which is closed under finite vector addition and scalar multiplication, with the property that

@@

Let $v_1,v_2,\ldots,v_k \in V$. A \textit{linear combination} of $v_1,v_2,\ldots,v_k \in V$ is any vector of the form

@@

A set of vectors $v_1,v_2,\ldots,v_k \in V$ is \textit{linearly independent} if the the only way to get

@@

Taken a set of linearly independent vectors $b = (v_1,\ldots,v_n) \in V$ we say that $b$ is a \textit{basis} of $V$ if $\forall w \in V$ we can write

@@

The vector's length or \textit{Euclidean norm} of $v = (x_1, x_2, \ldots, x_m)$ is

@@

Let $v, w \in V \subset \R^m$ and $v = (x_1, x_2, \ldots, x_m), w = (y_1, y_2, \ldots, y_m)$, the \textit{dot product} of $v$ and $m$ is

@@

\begin{figure}[!b]
    \centering
    \includegraphics[scale=0.2]{./img/dot_product.png}
    \caption{Dot Product By 3Blue1Brown \cite{3b1b}}
    \label{fig:dot_product}
\end{figure}

@@

An \textit{ortoghonal basis} for a vector space $V$ is a basis $v_1, \ldots, v_m$ with the property that

@@

If $\lVert v_i \rVert = 1$ for all $i$ then the basis is \textit{orthonormal}.

@@

Let $v_1,\ldots,v_n \in \R^m, m \ge n$ be linearly independent vectors. A \textit{Lattice} $L$ spanned by $\{v_1,\ldots,n_n\}$ is the set of 
all integer linear combinations of $v_1,\ldots,v_n$.

@@

If $v_i$ for every $i = 1,\ldots\,n$ has integer coordinates then the lattice is
called \textit{Integral Lattice}.

@@

\begin{algorithm}
    \caption{Gauss Basis Reduction}
    \vspace*{5px}
    \text{Loop} \\
    If $\lVert v_2 \rVert < \lVert v_1 \rVert$ , \text{swap }$v_1, v_2$ \\
    m = $\left \lceil{(v_1 \cdot v_2) {\lVert v_1 \rVert}^{-2}} \right \rceil$ \\
    \text{If } $m = 0$, \text{return } $v_1, v_2$ \\
    $v_2 = v_2 - mv_1$ \\
    \text{Continue Loop}
\end{algorithm}

@@

\chapter{RSA Cryptanalysis}

@@

\begin{itemize}
    \item $\approx$ 170 unkown bits of message from an RSA 1024-bit modulus
    \item $\approx$ 341 unkown bits of message from an RSA 2048-bit modulus
    \item $\approx$ 683 unkown bits of message from an RSA 4096-bit modulus
\end{itemize}

@@

\subsection{Factorize modulus with known MSB bits of $p$}

Knowing the most significat bits of $p$ can be used to factor the modulus $N$ ref TODO coppersmith.

Let $a = 2^lb$ where $b$ are the bits known of $p$, then $p = a + r$ for some small value of $r$.

Using as lattice basis

\[
    B = 
    \begin{pmatrix}
        X^2 & Xa & 0 \\
        0 & X & a \\
        0 & 0 & N
    \end{pmatrix} 
\]

And applying LLL we can find $r$.

\chapter{ECDSA Cryptanalysis}

@@

\subsection{Recover $d$ from many signatures with small $k_i$}

@@

\subsection{Recover $d$ knowing MSB bits of each $k_i$}
